Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-21 Master Thesis/synopsis.tex
Fix typos

\begin{proof}
\begin{enumerate}[(i)]
\item Consider the maximal unramified extension $L = K_P^\text{nr}$ of $K_P$ and let $w$ be the unique extension of $v_P$ in $L$. Now $k(w) = k(v_P)^\text{s}$. Since $w$ is unramified over $v$, $w$ is discrete as well. We have the following exact sequence:
\[ 1 \longrightarrow U_w/n \longrightarrow L^\times /n\longrightarrow \mathbb{Z}/n\mathbb{Z}\longrightarrow 0 \]
Since $k(w)$ is separably closed, with $\operatorname{char} k(w)=0$ and Hensel's lemma we can see that $U_w$ is a divisible group, hence we get an isomorphism $L^\times /n\cong \mathbb{Z}/n\mathbb{Z}$ induced by $w$. Therefore, the canonical map $K^\times/n\to L^\times/n$ can be identified with the map $K^\times/n \to\mathbb{Z}/n\mathbb{Z}$ induced by $v_P$. Now the following commutative diagram with exact columns gives us the result:
\[ \begin{tikzcd}
1 \ar[d] & 1 \ar[d]\\
U_{P'}/n \ar[r, dashed]\ar[d] & U_P/n\ar[d]\\
K'^\times /n \ar[r, "\varphi_n"]\ar[d] & K^\times/n\ar[d]\\
L'^\times /n \ar[r, "\varphi_n"'] &  L^\times /n
\end{tikzcd}\]
Note that $(K^\text{nr}_P)' = K'^{\text{nr}}_{P'}$ as proven in Lemma \ref{3.2}.
\item Consider the field extension $K_P/K$. Without loss of generality, we can assume $K=K_P$. Since $\operatorname{char}k(P)=0$, by Hensel's lemma, $r_P/n$ is an isomorphism and the composition
\[\begin{tikzcd}
k(P)^\times / n\ar[rr, "(r_P/n)^{-1}"] && U_P/n \ar[rr, hook] && K_P^\times/n
\end{tikzcd}\]
is the inflation $\operatorname{inf}: \mathrm{H}^1(k(P), \mu_n) \to \mathrm{H}^1(K_P,\mu_n)$. The same holds for $P'$ due to Lemma \ref{3.2}. Therefore, the assertion follows from the commutativity of the following diagram:
\[ \begin{tikzcd}
k(P')^\times/n \ar[r, "\cong"] \ar[d, "\varphi_P/n"'] & \mathrm{H}^1(k(P'),\mu_n) \ar[r, "\mathrm{inf}"] \ar[d]& \mathrm{H}^1(K'_{P'}, \mu_n) \ar[d, "\varphi_n"] \\
k(P)^\times/n \ar[r, "\cong"'] & \mathrm{H}^1(k(P),\mu_n)  \ar[r, "\mathrm{inf}"'] & \mathrm{H}^1(K_{P}, \mu_n)
\end{tikzcd}  \qedhere\]
\end{enumerate}
\end{proof}

@@

\begin{proof}
\begin{enumerate}[(i)]
\item By Corollary \ref{mordell-weil-corollary}, the Picard group of $X$ is finitely generated. Let $D_1,\ldots, D_r$ be Weil divisors of $X$ such that the associated line bundles generate $\operatorname{Pic}(X)$ and similarly let $D'_1,\ldots,D'_s$ generate $\operatorname{Pic}(X')$. Clearly any finite $S\subset X_0$ such that
\[ \bigcup_{i=1}^r \operatorname{supp}D_i\subset S\quad\text{and}\quad \bigcup_{j=1}^s\operatorname{supp}D'_j\subset S' \]
is admissible. Now let $V\subset X$ be an open subscheme. By the approximation theorem, we can choose the $D_i$ and the $D'_j$ in $V_0$ and $\sigma (V_0)$ respectively.
\item Observe that the following sequence is exact:
\[ \begin{tikzcd}
1 \ar[r] & K^S \ar[r]& K^\times \ar[r, "\operatorname{div}"]& \displaystyle \bigoplus_{P\in X_0\setminus S}\mathbb{Z} \ar[r]& \operatorname{Pic}(X\setminus S) = 0
\end{tikzcd} \]
Since $\bigoplus\mathbb{Z}$ is $\mathbb{Z}$-flat, tensoring with $\mathbb{Z}/n\mathbb{Z}$ yields another exact sequence and we can consider $K^S/n$ as a subgroup of $K^\times/n$. Then:
\[ K^S/n = \bigcap_{P\in X_0\setminus S}U_P/n \]
The analogous result holds for $K'^{S'}/n$. Thus the assertion follows from Lemma \ref{4.3} (i).\qedhere
\end{enumerate}
\end{proof}

@@

\begin{proof}
Let $U = X\setminus S$ and $P_1\in U$ be a closed point. We set:
\[ C_1 = \ker(r_{P_1}|_{K^S}: K^S \to k(P_1)^\times) \]
Since $r_{P_1}|_{k^\times}:k^\times \to k(P_1)^\times$ is injective, we have $C_1\cap k^\times = 1$ and the following map is injective:
\[ C_1 \to\prod_{P\in S}\mathbb{Z},\ x \mapsto (v_P(x))_{P\in S} \]
Therefore, $C_1$ is a free abelian group of finite rank. If $C_1 = 1$, then we set $T = \{P_1\}$ and we are done. Otherwise, let $x\in C_1,\ x\neq 1$. Then there exists an $a\in k^\times$ with the following properties:
\begin{enumerate}[(i)]
\item There exists a point $P_2\in X_0\setminus(S\cup\{P_1\})$ which lies above the prime of $k(x)$ corresponding to the polynomial $x-a\in k[x]$.
\item $a$ is not a root of unity.
\end{enumerate}
We set $C_2 = \ker(r_{P_2}|_{C_1})\subset C_1$. $C_2$ is free abelian of finite rank as well. By requirement $r_{P_2}(x)=a$ and $a\in k^\times$ is not of finite order, thus $x\not\in C_2$. It follows $\operatorname{rank}(C_2) < \operatorname{rank}(C_1)$. By repeating this process until $C_r=1$, we get points $P_2,\ldots,P_r\in U_0$, such that the following map is injective:
\[ C_1\to\prod_{i=2}^{r} k(P_i),\ x\mapsto(r_{P_i}(x))_i \]
Finally for $T=\{P_1,\ldots,P_r\}\subset U_0$, the map $r_{S,T}$ is injective. For the second assertion, note the following isomorphism with the snake lemma:
\[ \operatorname{coker}(r_{S,T}) \cong\operatorname{coker}\Big(K^S/k^\times \to \prod_{P\in T}k(P)^\times /k^\times \Big) \]
Since $K^S/k^\times$ is finitely generated, Lemma \ref{5.2} (iii) and Lemma \ref{5.3} (ii) show that the latter group is free-by-finite.
\end{proof}

@@

\begin{proof}
By Lemma \ref{5.4}, the following maps are injective for all admissible subsets $S\subset X_0$:
\[ r_S:K^S \to\prod_{P\in X_0\setminus S} k(P)^\times,\ x\mapsto (r_P(x))_P \]
Therefore, $\varphi$ is unique. For existence, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the composition $\varrho:K'^{S'}\to\prod k(\sigma P)^\times\to \prod k(P)^\times\to\operatorname{coker}(r_{S,T})$ is trivial. Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.

>>>>

\begin{proof}
\begin{enumerate}[(i)]
\item Consider the maximal unramified extension $L = K_P^\text{nr}$ of $K_P$ and let $w$ be the unique extension of $v_P$ in $L$. Now $k(w) = k(v_P)^\text{s}$. Since $w$ is unramified over $v$, $w$ is discrete as well. We have the following exact sequence:
\[ 1 \longrightarrow U_w/n \longrightarrow L^\times /n\longrightarrow \mathbb{Z}/n\mathbb{Z}\longrightarrow 0 \]
Since $k(w)$ is separably closed, with $\operatorname{char} k(w)=0$ and Hensel's lemma we can see that $U_w$ is a divisible group, hence we get an isomorphism $L^\times /n\cong \mathbb{Z}/n\mathbb{Z}$ induced by $w$. Therefore, the canonical map $K^\times/n\to L^\times/n$ can be identified with the map $K^\times/n \to\mathbb{Z}/n\mathbb{Z}$ induced by $v_P$. Now the following commutative diagram with exact columns gives us the result:
\[ \begin{tikzcd}
1 \ar[d] & 1 \ar[d]\\
U_{P'}/n \ar[r, dashed]\ar[d] & U_P/n\ar[d]\\
K'^\times /n \ar[r, "\varphi_n"]\ar[d] & K^\times/n\ar[d]\\
L'^\times /n \ar[r, "\varphi_n"'] &  L^\times /n
\end{tikzcd}\]
Note that $(K^\text{nr}_P)' = K'^{\text{nr}}_{P'}$ as proven in Lemma \ref{3.2}.
\item Consider the field extension $K_P/K$. Without loss of generality, we can assume $K=K_P$. Since $\operatorname{char}k(P)=0$, by Hensel's lemma, $r_P/n$ is an isomorphism and the composition
\[\begin{tikzcd}
k(P)^\times / n\ar[rr, "(r_P/n)^{-1}"] && U_P/n \ar[rr, hook] && K_P^\times/n
\end{tikzcd}\]
coincides with the inflation $\operatorname{inf}: \mathrm{H}^1(k(P), \mu_n) \to \mathrm{H}^1(K_P,\mu_n)$. The same holds for $P'$ due to Lemma \ref{3.2}. Therefore, the assertion follows from the commutativity of the following diagram:
\[ \begin{tikzcd}
k(P')^\times/n \ar[r, "\cong"] \ar[d, "\varphi_P/n"'] & \mathrm{H}^1(k(P'),\mu_n) \ar[r, "\mathrm{inf}"] \ar[d]& \mathrm{H}^1(K'_{P'}, \mu_n) \ar[d, "\varphi_n"] \\
k(P)^\times/n \ar[r, "\cong"'] & \mathrm{H}^1(k(P),\mu_n)  \ar[r, "\mathrm{inf}"'] & \mathrm{H}^1(K_{P}, \mu_n)
\end{tikzcd}  \qedhere\]
\end{enumerate}
\end{proof}

@@

\begin{proof}
\begin{enumerate}[(i)]
\item By Corollary \ref{mordell-weil-corollary}, the Picard group of $X$ is finitely generated. Let $D_1,\ldots, D_r$ be Weil divisors of $X$ such that the associated line bundles generate $\operatorname{Pic}(X)$ and similarly let $D'_1,\ldots,D'_s$ generate $\operatorname{Pic}(X')$. Clearly any finite $S\subset X_0$ such that
\[ \bigcup_{i=1}^r \operatorname{supp}D_i\subset S\quad\text{and}\quad \bigcup_{j=1}^s\operatorname{supp}D'_j\subset S' \]
is admissible. Now let $V\subset X$ be an open subscheme. By the approximation theorem, we can choose the $D_i$ and the $D'_j$ so that their support is in $V_0$ and $\sigma (V_0)$ respectively.
\item Observe that the following sequence is exact:
\[ \begin{tikzcd}
1 \ar[r] & K^S \ar[r]& K^\times \ar[r, "\operatorname{div}"]& \displaystyle \bigoplus_{P\in X_0\setminus S}\mathbb{Z} \ar[r]& \operatorname{Pic}(X\setminus S) = 0
\end{tikzcd} \]
Since $\bigoplus\mathbb{Z}$ is $\mathbb{Z}$-flat, tensoring with $\mathbb{Z}/n\mathbb{Z}$ yields another exact sequence and we can consider $K^S/n$ as a subgroup of $K^\times/n$. Then:
\[ K^S/n = \bigcap_{P\in X_0\setminus S}U_P/n \]
The analogous result holds for $K'^{S'}/n$. Thus the assertion follows from Lemma \ref{4.3} (i).\qedhere
\end{enumerate}
\end{proof}

@@

\begin{proof}
Let $U = X\setminus S$ and $P_1\in U$ be a closed point. We set:
\[ C_1 = \ker(r_{P_1}|_{K^S}: K^S \to k(P_1)^\times) \]
Since $r_{P_1}|_{k^\times}:k^\times \to k(P_1)^\times$ is injective, we have $C_1\cap k^\times = 1$ and the following map is injective:
\[ C_1 \to\prod_{P\in S}\mathbb{Z},\ x \mapsto (v_P(x))_{P\in S} \]
Therefore, $C_1$ is a free abelian group of finite rank. If $C_1 = 1$, then we set $T = \{P_1\}$ and we are done. Otherwise, let $x\in C_1,\ x\neq 1$. Then there exists an $a\in k^\times$ with the following properties:
\begin{enumerate}[(i)]
\item There exists a point $P_2\in X_0\setminus(S\cup\{P_1\})$ which lies above the prime of $k(x)$ corresponding to the polynomial $x-a\in k[x]$.
\item $a$ is not a root of unity.
\end{enumerate}
We set $C_2 = \ker(r_{P_2}|_{C_1})\subset C_1$. $C_2$ is free abelian of finite rank as well. By requirement $r_{P_2}(x)=a$ and $a\in k^\times$ is not of finite order, thus $x\not\in C_2$. It follows $\operatorname{rank}(C_2) < \operatorname{rank}(C_1)$. By repeating this process until $C_r=1$, we get points $P_2,\ldots,P_r\in U_0$, such that the following map is injective:
\[ C_1\to\prod_{i=2}^{r} k(P_i),\ x\mapsto(r_{P_i}(x))_i \]
Finally for $T=\{P_1,\ldots,P_r\}\subset U_0$, the map $r_{S,T}$ is injective. For the second assertion, note the following isomorphism:
\[ \operatorname{coker}(r_{S,T}) \cong\operatorname{coker}\Big(K^S/k^\times \to \Big(\prod_{P\in T}k(P)^\times \Big)\Big/k^\times \Big) \]
Since $K^S/k^\times$ is finitely generated, Lemma \ref{5.2} (iii) and Lemma \ref{5.3} (ii) show that the latter group is free-by-finite.
\end{proof}

@@

\begin{proof}
By Lemma \ref{5.4}, the following maps are injective for all admissible subsets $S\subset X_0$:
\[ r_S:K^S \to\prod_{P\in X_0\setminus S} k(P)^\times,\ x\mapsto (r_P(x))_P \]
Now it's evident that $\varphi$ is unique. For existence, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the composition $\varrho:K'^{S'}\to\prod k(\sigma P)^\times\to \prod k(P)^\times\to\operatorname{coker}(r_{S,T})$ is trivial. Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.
